### 2022-06-20 appendixes/0A2-appendix-2.tex
txs auto checkin

\chapter{Lampiran B. Tangkapan Layar Hasil Pengujian Kinerja}