### 2022-06-20 appendixes/0A2-appendix-2.tex
txs auto checkin

\chapter{Lampiran B. Tangkapan Layar Hasil Pengujian Kinerja}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.5\textwidth]{resources/ch4/resource/1-cpu.png}
	\caption{Penggunaan CPU pada kasus \textbf{SI1}}
	\label{result_cpu_1}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.5\textwidth]{resources/ch4/resource/1-mem.png}
	\caption{Penggunaan Memory pada kasus \textbf{SI1}}
	\label{result_mem_1}
\end{figure}